\MakeUppercase{Graph Product Structure for Non-Minor-Closed Classes}

### 2022-03-19 k-planar.tex
Fixed mismatched braces
Pfffft

All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this statement.} These applications motivate finding product structure theorems for other graph classes. \citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of a surface with $h$ handles and $c$ crosscaps is $2h+c$. The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$. Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} and for apex-minor-free graphs\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges. A class $\mathcal{G}$ of graphs is \defin{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$. A minor-closed class is \defin{proper} if it is not the class of all graphs. For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class. A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar. A 1-apex graph is \defin{apex}.  A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.}, and \citet{DEMWW22} do so for graphs in any minor-closed class and with bounded maximum degree. See \cref{Generalisations} for more precise statements and see \citep{DHJLW21} for a survey on this topic.

@@

A \defin{layering} of a graph $G$ is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$ 
\note{DW}{just write ``is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge ...''}
and for every edge $vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.  For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$. To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the vertex it creates in $H$ \note{DW}{``is indexed by the corresponding vertex in $H$''.}

@@

As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
\citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cite{UWY}). The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering. At the core of their work is the elegant proof by \citet{PS21} of the following result:

@@

Consider a tree $T$ rooted at some node $x_0\in V(T)$. Each subtree $T'$ of $T$ is considered to be rooted at the node in $T'$ closest to $x_0$. A node $a\in V(T)$ is a \defin{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \defin{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.  Note that each node $x\in V(T)$ is a $T$-ancestor and $T$-descendant of itself.  We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \defin{strict} $T$-ancestor of $x$ if $a\neq x$.
The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.
% For each node $x\in V(T)$, define
% \[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
% to be the maximal subtree of $T$ rooted at $x$.

@@

  Begin with any width-$t$ tree-decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$ and has no empty bags.  Select any node $x\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x$ and root $T_0$ at $x_0$. (The purpose of $x_0$ is to ensure that every node $x$ for which $B_x$ is non-empty has a parent.)  Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.  If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.  Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$. Now $f(y)=a'$ and $f(x)=a$, so this modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$, so repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.

  Next, remove the node $x_0$ from $T_0$, and consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.  In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.  In this case, contract the edge $aa'$ in $T_0$, eliminating the node $a$.  Repeating this operation will eventually produce a width-$t$ tree-decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.  Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.  By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
\end{proof}

@@

  Suppose, without loss of generality, that $v_0$ participates in $x_i$.  Then $a(v_0)=x$ is a $T$-ancestor of $x_i$ which a $T$-ancestor of $x_r$.  Finally $a(v_r)=y$ is a $T$-ancestor of $x_r$.  Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so one of $x$ or $y$ is a $T$-ancestor of the other.
\end{proof}

@@

  It remains to show that each bag in our tree decomposition has size at most $\binom{k+t}{t}$.  We will do this by appealing to an elegant result of \citet{PS21} on $k$-reachability in directed skeletons, which we now explain.

  Let $H^+$ denote the supergraph of $H$ that contains the edge $xy$ if and only if some bag $B_z\in\mathcal{T}$ contains both $x$ and $y$ (i.e., $H^+$ is the chordal closure of $H$ with respect to $\mathcal{T}$).  Let $\overrightarrow{H}^+$ be the directed graph obtained by directing each edge $xy$ of $H^+$ in the direction $\overrightarrow{xy}$ so that $y$ is a $T$-ancestor of $x$. Observe that directed path in $\overrightarrow{H}^+$ that begins at $x$ leads to a $T$-ancestor of $x$.  \citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of strict $T$-ancestors of $x$ that can be reached by a directed path of length at most $k$ that begins at $x$ is at most $\binom{k+t}{t}-1$.

@@

  Therefore $W:=y_0,\ldots,y_s:=x,x_j,x_{j+1},\ldots,x_r$ is a lazy walk in $H^+$ of length $r\le k$.  For each $j\in \{0,\ldots,s\}$, let $z_j$ be the lowest common $T$-ancestor of $y_0,\ldots,y_j$.  The same reasoning used to conclude that $xx_j\in E(H^+)$ can be used to conclude that $z_{i-1}=z_i$ or $z_{i-1}z_i\in E(H^+)$ for each $i\in\{1,\ldots,s\}$.  Therefore $z_0,\ldots,z_s$ is a directed lazy walk in $\overrightarrow{H}^+$.  Finally, let $\overrightarrow{W}$ be the path in $\overrightarrow{H}^+$ obtained by removing removing duplicates from $z_0,\ldots,z_s$.  Then $\overrightarrow{W}$ is a directed path in $\overrightarrow{H}^+$ from $x$ to $a$ of length $s\le r\le k$.  This completes the proof.
\end{proof}

At this point, the proof of \cref{ShortcutPartition} is almost immediate from \cref{general-width,general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\SS}$.  In particular, $G^{\SS}$ may contain edges $vw$ with $v\in L_i$ and $w\in L_{i+j}$ for any $j\in\{0,\ldots,k\}$.  To resolve this, we use a new layering $\mathcal{L}':=\langle L_0',\ldots,L_h'\rangle$ in which $L_i'=\bigcup_{j=ki}^{ki+k-1} L_i$.  This increases the layered width given by \cref{general-width} from $d\ell(k^2+3)$ to $d\ell(k^3+3k)$.  Therefore $G$ has an $H$-partition of layered width at most $d\ell(k^3+3k)$ in which $H$ has treewidth at most $\binom{k+t}{t}$, completing the proof of \cref{ShortcutPartition}.
\end{proof}

@@

@@

@@

\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
		\item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
		\item Let $F$ be a cycle in $G'$ that $V(F)$ is partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
		\begin{compactenum}
			\item $F[P_i]$ is connected; and
			\item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
			\note{DW}{suggested revision: ``$P_i$ has a partition $\{X_i,Y_i\}$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.''}
		\end{compactenum}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
	\begin{compactenum}[(i)]
		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
		\item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$ \note{DW}{suggested revision, ``$S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$''};
		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
	\end{compactenum}
%\note{PM}{I've changed my mind about this notation change.  We end up applying the lemma inductively on the faces $F_1,\ldots,F_m$ of $M$ (which are cycles in $G$).  It makes sense that the original cycle in $G$ is called $F$.}
%\note{DW}{okay}
%\begin{lem}
%  \label{induction} The setup:
%  \begin{compactenum}
%    \item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
%    \item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
%    \item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
%    \item Let $F$ be a cycle in $G'$ whose vertices are partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
%    \begin{compactenum}
%      \item $F[P_i]$ is connected; and
%      \item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
%    \end{compactenum}
%    \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
%  \end{compactenum}
%  Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
%  \begin{compactenum}[(i)]
%    \item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
%    \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
%    \item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
%  \end{compactenum}
\end{lem}

@@

	The proof is by induction on the number of inner faces of $N'$.  A particular easy case occurs when $V(N')=V(F)$, for which we can simply take 
	\note{DW}{Suggested revision, ``In the base case, $N'$ has only one inner face, and we can simply take ...} 	
	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
	\begin{compactenum}[(i)]
		\item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
		\item Assumption 4(b) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
	\end{compactenum}

@@

In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.  We first explain how $v_1$, $v_2$ and $v_3$ are chosen and explain how these are used to define $S$.  There is one case to consider for each possible value of $k$ (see \cref{boring_figure}):

@@

		\item If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the appropriate vertex of $F_0$.  (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.  Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add edges $d'-2$
		\note{DW}{should be $d'-3$, right?} edges inside of $F$ to split it into $3$-sided faces.  This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary. By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.  Let $\tau:=v_1,\ldots,v_p$ be the face \note{DW}{``be the facial cycle''} of $N'$ that contains $\tau''$.		
	\end{compactenum}

@@

\cref{NewMapGraph} is an immediate consequence of \cref{d_framed_product_stucture}} and the next lemma. Similar results to \cref{MewMapGraphLemma} appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.

>>>>

All of these results hold for any graph class that has a product structure theorem analogous to \cref{PlanarProduct}; that is, for any graph class  $\mathcal{G}$ where every graph in $\mathcal{G}$ is contained in $H\boxtimes P\boxtimes K_\ell$ where $H$ has bounded treewidth, $P$ is a path, and $\ell$ is bounded.\footnote{It is easily seen that $\tw(H\boxtimes K_\ell) \leq (\tw(H)+1)\ell-1$, so we may assume that $\ell=1$ in this statement.} These applications motivate finding product structure theorems for other graph classes. \citet{DJMMUW20} prove product structure theorems for graphs of bounded Euler genus\footnote{The \textit{Euler genus} of a surface with $h$ handles and $c$ crosscaps is $2h+c$. The \textit{Euler genus} of a graph $G$ is the minimum integer $g$ such that $G$ embeds in a surface of Euler genus $g$. Of course, a graph is planar if and only if it has Euler genus 0; see \citep{mohar.thomassen:graphs} for more about graph embeddings in surfaces.} and for apex-minor-free graphs,\footnote{A graph $M$ is a \textit{minor} of a graph $G$ if a graph isomorphic to $M$ can be obtained from a subgraph of $G$ by contracting edges. A class $\mathcal{G}$ of graphs is \defin{minor-closed} if for every graph $G\in\mathcal{G}$, every minor of $G$ is in $\mathcal{G}$. A minor-closed class is \defin{proper} if it is not the class of all graphs. For example, for fixed $g\geq 0$, the class of graphs with Euler genus at most $g$ is a proper minor-closed class. A graph $G$ is $t$-apex if it contains a set $A$ of at most $t$ vertices such that $G-A$ is planar. A 1-apex graph is \defin{apex}.  A minor-closed class $\mathcal{G}$ is apex-minor-free if some apex graph is not in $\mathcal{G}$.} and \citet{DEMWW22} do so for graphs in any minor-closed class and with bounded maximum degree. See \cref{Generalisations} for more precise statements and see \citep{DHJLW21} for a survey on this topic.

@@

A \defin{layering} of a graph $G$
% is a sequence $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that $\{L_0,L_1,\ldots\}$ is a partition of $V(G)$
% \note{DW}{just write ``
is an ordered partition $\mathcal{L}=\langle L_0,L_1,\ldots\rangle $ such that for every edge
 % ...''}
$vw\in E(G)$, if $v\in L_i$ and $w\in L_j$ then $|j-i|\leq 1$.  For any partition $\PP=\{S_1,\ldots,S_p\}$ of $V(G)$, a \defin{quotient graph} $H=G/\PP$ has a $p$-element vertex set $V(H)=\{x_1,\ldots,x_p\}$ and $x_ix_j\in E(H)$ if and only if there exists an edge $vw\in E(G)$ such that $v\in S_i$ and $w\in S_j$. To highlight the importance of the quotient graph $H$, we call $\PP$ an \defin{$H$-partition} and write this concisely as $\PP=\{S_x : x\in V(H)\}$ so that each element of $\PP$ is indexed by the corresponding vertex in $H$.

@@

As an example of the use of layered partitions, to prove \cref{PlanarProduct}(a),
\citet{DJMMUW20} showed that every planar graph has an $H$-partition of layered width $1$ for some planar graph $H$ of treewidth at most $8$ (improved to $6$ in \cite{UWY}). The proof is constructive and gives a simple quadratic-time algorithm for finding the corresponding partition and layering.\footnote{\citet{bose.morin.ea:optimal} have recently given linear time algorithms for computing layered $H$-partitions of planar graphs.} At the core of their work is the elegant proof by \citet{PS21} of the following result:

@@

Consider a tree $T$ rooted at some node $x_0\in V(T)$. Each subtree $T'$ of $T$ is considered to be rooted at the node in $T'$ closest to $x_0$. A node $a\in V(T)$ is a \defin{$T$-ancestor} of $x\in V(T)$ (and $x$ is a \defin{$T$-descendant} of $a$) if $a$ is a vertex of the path, in $T$, from $x_0$ to $x$.  Note that each node $x\in V(T)$ is a $T$-ancestor and $T$-descendant of itself.  We say that a $T$-ancestor $a\in V(T)$ of $x\in V(T)$ is a \defin{strict} $T$-ancestor of $x$ if $a\neq x$.
The \defin{$T$-depth} of a node $x\in V(T)$ is the length of the path, in $T$, from $x_0$ to $x$.  The \defin{lowest common ancestor} of a non-empty set $S\subseteq V(T)$ is the maximum $T$-depth node of $T$ that is a $T$-ancestor of every node in $S$.
% For each node $x\in V(T)$, define
% \[T_x := T[\{y\in V(T):\mbox{$x$ is a $T$-ancestor of $y$}\}] \]
% to be the maximal subtree of $T$ rooted at $x$.

@@

  Begin with any width-$t$ tree-decomposition $(B_x:x\in V(T_0))$ of $H$ that uses some tree $T_0$ and has no empty bags.  Select any node $x_1\in V(T_0)$, add a leaf $x_0$, with $B_{x_0}=\emptyset$, adjacent to $x_1$ and root $T_0$ at $x_0$. (The purpose of $x_0$ is to ensure that every node $x$ for which $B_x$ is non-empty has a parent.)  Let $f:V(H)\to V(T)$ be the function that maps each $x\in V(H)$ onto the root of the subtree $T_0[x]:=T_0[\{y\in V(T_0): x\in B_y\}]$.  If $f$ is not one-to-one, then select some distinct pair $x,y\in V(H)$ with $a:=f(x)=f(y)$.  Subdivide the edge between $a$ and its parent in $T_0$ by introducing a new node $a'$ with $B_{a'}=B_{a}\setminus\{x\}$. Now $f(y)=a'$ and $f(x)=a$, so this modification reduces the number of distinct pairs $x,y\in V(H)$ with $f(x)=f(y)$.  Repeatedly performing this modification will eventually produce a tree-decomposition $(B_x:x\in V(T_0))$ of $H$ in which $f$ is one-to-one.

  Next, remove the node $x_0$ from $T_0$ (so that $x_1$ becomes the new root of $T_0$).  Consider any node $a\in V(T_0)$ such that there is no vertex $x\in V(H)$ with $f(x)=a$.  In this case, $B_{a}\subseteq B_{a'}$ where $a'$ is the parent of $a$ since any $x\in B_a\setminus B_{a'}$ would have $f(x)=a$.  In this case, contract the edge $aa'$ in $T_0$, eliminating the node $a$.  Repeating this operation will eventually produce a width-$t$ tree-decomposition of $(B_x:x\in V(T_0))$ where $f$ is a bijection between $V(H)$ and $V(T_0)$.  Renaming each node $a\in V(T_0)$ as $f^{-1}(a)$ gives a tree-decomposition $(B_x:x\in V(T))$ with $V(T)=V(H)$.  By the definition of $f$, the tree-decomposition $(B_x:x\in V(T))$ satisfies \tref{subtree-root}.
\end{proof}

@@

  Suppose, without loss of generality, that $v_0$ participates in $x_i$.  Then $a(v_0)=x$ is a $T$-ancestor of $x_i$ which a $T$-ancestor of $x_r$.  Finally $a(v_r)=y$ is a $T$-ancestor of $x_r$.  Therefore, both $x$ and $y$ are contained in the path from the root of $T$ to $x_r$, so at least one of $x$ or $y$ is a $T$-ancestor of the other.
\end{proof}

@@

  It remains to show that each bag in this tree decomposition has size at most $\binom{k+t}{t}$.  We will do this by appealing to an elegant result of \citet{PS21} on $k$-reachability in directed skeletons, which we now explain.

  Let $H^+$ denote the supergraph of $H$ that contains the edge $xy$ if and only if some bag $B_z\in\mathcal{T}$ contains both $x$ and $y$ (i.e., $H^+$ is the chordal closure of $H$ with respect to $\mathcal{T}$).  Let $\overrightarrow{H}^+$ be the directed graph obtained by directing each edge $xy$ of $H^+$ in the direction $\overrightarrow{xy}$ so that $y$ is a $T$-ancestor of $x$. Observe that any directed path in $\overrightarrow{H}^+$ that begins at $x$ leads to a $T$-ancestor of $x$.  \citet[Lemma~13]{PS21} show that, for any $x\in V(\overrightarrow{H}^+)$, the number of strict $T$-ancestors of $x$ that can be reached by a directed path of length at most $k$ that begins at $x$ is at most $\binom{k+t}{t}-1$.

@@

  Therefore $W:=y_0,\ldots,y_s:=x,x_j,x_{j+1},\ldots,x_r$ is a lazy walk in $H^+$ of length $r\le k$.  For each $j\in \{0,\ldots,s\}$, let $z_j$ be the lowest common $T$-ancestor of $y_0,\ldots,y_j$.  The same reasoning used to conclude that $xx_j\in E(H^+)$ can be used to conclude that $z_{i-1}=z_i$ or $z_{i-1}z_i\in E(H^+)$ for each $i\in\{1,\ldots,s\}$.  Therefore $z_0,\ldots,z_s$ is a directed lazy walk in $\overrightarrow{H}^+$.  Finally, let $\overrightarrow{W}$ be the path in $\overrightarrow{H}^+$ obtained by removing removing duplicates from $z_0,\ldots,z_s$.  Then $\overrightarrow{W}$ is a directed path in $\overrightarrow{H}^+$ from $x$ to $a$ of length at most $s\le r\le k$.  This completes the proof.
\end{proof}

At this point, the proof of \cref{ShortcutPartition} is almost immediate from \cref{general-width,general-bag-size}, except that the layering $\mathcal{L}$ of $G$ may not be a valid layering of $G^{\SS}$.  In particular, $G^{\SS}$ may contain edges $vw$ with $v\in L_i$ and $w\in L_{i+j}$ for any $j\in\{0,\ldots,k\}$.  To resolve this, we use a new layering $\mathcal{L}':=\langle L_0',\ldots,L_h'\rangle$ in which $L_i'=\bigcup_{j=ki}^{ki+k-1} L_i$.  This increases the layered width given by \cref{general-width} from $d\ell(k^2+3)$ to $d\ell(k^3+3k)$.  Therefore $G$ has an $H$-partition of layered width at most $d\ell(k^3+3k)$ in which $H$ has treewidth at most $\binom{k+t}{t}-1$, completing the proof of \cref{ShortcutPartition}.
\end{proof}

@@

\note{PM}{I don't like this structure anymore.  The proof of \cref{kPlanarProduct} ($k$-planar) refers to the innards of the proof of \cref{ShortcutPartition} ($(k,d)$-shortcuts).  The proofs should be close to each other.

Let's consider something like this:
\begin{compactenum}
  \item Introduction
  \item Shortcut systems and $k$-planar graphs
  \item All graphs obtained from shortcut systems or $k$-planar graphs
  \item $d$-framed graphs, $1$-planar graphs, and map graphs
  \item All applications (colouring, labelling, etc).
\end{compactenum}
% (Possibly with the order of 2 and 3 reversed.)
}

@@

\note{PM}{I would also like to change $G'$ to $G_0$ and $N'$ to $N_0$.  One issue is holding me back, though: We recursive on cycles $F_1,\ldots,F_m$ that define subgraphs $N_1,\ldots,N_m$ of $G$ and subgraphs $N_1',\ldots,N_m'$ of $G'$.  Will these latter subgraphs becomes $N_{0,1},\ldots,N_{0,m}$?}

@@

\begin{lem}
	\label{induction} The setup:
	\begin{compactenum}
		\item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
		\item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
		\item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
		\item Let $F$ be a cycle in $G'$ that $V(F)$ is partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
		\begin{compactenum}
			\item $F[P_i]$ is connected; and
			\item $P_i$ has a partition $\{X_i,Y_i\}$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
		\end{compactenum}
		\item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the interior or boundary of $F$.
	\end{compactenum}
	Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
	\begin{compactenum}[(i)]
		\item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
		\item For each $x\in V(H)$, $S_x$ has a partition $\{X_x,Y_x\}$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
		\item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
	\end{compactenum}
%\note{PM}{I've changed my mind about this notation change.  We end up applying the lemma inductively on the faces $F_1,\ldots,F_m$ of $M$ (which are cycles in $G$).  It makes sense that the original cycle in $G$ is called $F$.}
%\note{DW}{okay}
%\begin{lem}
%  \label{induction} The setup:
%  \begin{compactenum}
%    \item Let $G$ be a $d$-framed multigraph and let $G'$ be a frame of $G$.
%    \item Let $F_0$ be the outer face of $G'$ and let $T$ be a BFS spanning forest of $G'$ rooted at $V(F_0)$.
%    \item For every integer $j\ge 0$, let $L_j=\{v\in V(G):\dist_T(v,V(F_0))=j\}$.
%    \item Let $F$ be a cycle in $G'$ whose vertices are partitioned into $k\le 3$ sets $P_1,\ldots,P_k$ such that for each $i\in\{1,\ldots,k\}$,
%    \begin{compactenum}
%      \item $F[P_i]$ is connected; and
%      \item $P_i=Y_i\cup X_i$ where $|X_i|\le d-3$ and $|Y_i\cap L_j| \le 3$ for each $j\ge 0$.
%    \end{compactenum}
%    \item Let $N$ and $N'$ be the subgraphs of $G$ and $G'$ consisting only of those edges and vertices contained in the closure of $F$.
%  \end{compactenum}
%  Then $N$ has an $H$-partition $\PP=\{S_x: x\in V(H)\}$ such that:
%  \begin{compactenum}[(i)]
%    \item for each $i\in\{1,\ldots,k\}$, there exists $x_i\in V(H)$ such that $P_i=S_{x_i}$;
%    \item For each $x\in V(H)$, $S_x=Y_x\cup X_x$ where $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each $j\ge 0$;
%    \item $H$ has a $3$-simple tree decomposition $\mathcal{T}$ and, if $k=3$, then $\{x_1,x_2,x_3\}$ is contained in exactly one bag of $\mathcal{T}$.
%  \end{compactenum}
\end{lem}

@@

	The proof is by induction on the number of inner faces of $N'$.  A particular easy case occurs when $V(N')=V(F)$, for which we can simply take
	\note{DW}{Suggested revision, ``In the base case, $N'$ has only one inner face, and we can simply take ...} \note{PM}{I debated what to put here since, even when $N'$ has only one face, the rules below work for picking $\tau$.  The \emph{real} base case occurs when $N'$ has zero faces, but that's not something that ever happens since we only ever apply induction on the inner faces of $M$ (except $\tau$).  I settled on this ``base case'' since it's clear that, in this case, the partition $\mathcal{P}$ is already decided.}
	$\PP:=\{P_1,\ldots,P_k\}$ and verify that the preconditions of the lemma (the `setup') ensure that $\PP$ satisfies the requirements (i)--(iv):
	\begin{compactenum}[(i)]
		\item By definition, each of $P_1,\ldots,P_k$ is a part of $\mathcal{P}$.
		\item Assumption 4(b) ensures, for each $x\in V(H)$, the existence of $X_x$ and $Y_x$ such that $|X_x|\le d-3$ and $|Y_x\cap L_j|\le 3$ for each integer $j\ge 0$.
		\item The trivial tree decomposition of $H$ that has a single bag containing $\{x_1,\ldots,x_k\}$ is $3$-simple and, if $k=3$, has exactly one bag that contains $\{x_1,\ldots,x_3\}$.
	\end{compactenum}

@@

In the general case, our $H$-partition $\mathcal{P}$ will include $P_1$, $P_2$ , $P_3$ and a (possibly empty) part $S$ that will be determined by three vertices $v_1v_2v_3$ that belong to a common inner face $\tau$ of $N'$.  We first explain how $v_1$, $v_2$ and $v_3$ are chosen and then explain how these are used to define $S$.  There is one case to consider for each possible value of $k$ (see \cref{boring_figure}):

@@

		\item If $k=3$ then assign each vertex $v$ of $N'$ a colour $\alpha(v)\in\{1,\ldots,3\}$ as follows:  Let $w$ be the first vertex of $F$ encountered on the path in $T$ from $v$ to the appropriate vertex of $F_0$.  (This implies that $w=v$ when $v$ is a vertex of $F$.)  Since $w\in V(F)$, $w\in P_i$ for some $i\in\{1,\ldots,k\}$ and we define $\alpha(v):=i$.  Now, for each $d'\in\{4,\ldots,d\}$ and each $d'$-sided inner face $F$ of $N'$, add $d'-3$ edges inside of $F$ to split it into $3$-sided faces.  This yields a plane supergraph $N''$ of $N'$ in which each inner face has exactly three edges on its boundary. By Sperner's Lemma there exists an inner face $\tau'':=v_1,v_2,v_3$ of $N''$ such that $\alpha(v_i)=i$ for each $i\in\{1,2,3\}$.  Let $\tau:=v_1,\ldots,v_p$ be  the facial cycle of $N'$ that contains $\tau''$.
	\end{compactenum}

@@

\cref{NewMapGraph} is an immediate consequence of \cref{d_framed_product_stucture} and the next lemma. Similar results to \cref{MewMapGraphLemma} appear in earlier works \cite{CGP06,BDGGMR,Brandenburg19,Brandenburg20} but we state the precise lemma and provide a proof here for the sake of completeness.
